### 2024-11-07 Data_Cleaning_20241031.Rmd
Final Version of Draft Exploratory Analysis
Fixed Y label for license over time graph

---
title: "ECNS 560 Project - Bison Population"
author: "Kyle Froisland"
date: "2024-10-16"
output: html_document
---

Load Raw Data/Packages: *Data loaded from path on my local computer
```{r}
library(readr)
library(tidyverse)
library(readxl)
library(sf)
library(tmap)
library(raster)
library(rvest)
library(units)
library(dplyr)
library(ggplot2)

setwd("C:/Users/kdfro/Documents/MSU/ECNS 560/Term Project")
Bison_numbers <- read_csv("Data/Bison_numbers_by_the_county National Bison Association.csv")
hunting_data_tidy <- read_excel("Data/hunting_data_tidy.xlsx")
```

Create Tidy Version of Population Data:
```{r}
tidy_bison = Bison_numbers |>
  pivot_longer(cols=bison_census2002:bison_census2017, names_to="year", values_to="population") |>
  dplyr::select(state_code, county, year, population)
```

Cleaning:
```{r}
#Change year Variable in tidy_bison to just the number (remove "bison_census")
tidy_bison_clean = mutate(tidy_bison, year = sapply(str_remove(tidy_bison$year, "bison_census"), function(x) (x[1])))

#define new dataset with by-state population vs by county
bison_data_merge = tidy_bison_clean |>
  group_by(state_code, year) |>
  summarize(total_population = sum(population)) |>
  rename_at(2, ~'year')

#Change column name of year in hunting data, filter to only years that match population data
colnames(hunting_data_tidy)[2] = 'year' 
hunting_data_tidy |>
  filter(year == "2002" | year == "2007" | year == "2012" | year == "2017")

#Ensure that years are numeric
bison_data_merge$year = as.numeric(bison_data_merge$year)
hunting_data_tidy$year = as.numeric(hunting_data_tidy$year)

```

Merge Data Sets:
```{r}
#join with hunting data by state and year as ID
merged_bison_data = left_join(bison_data_merge, hunting_data_tidy, by = c("state_code" = "state_code", "year" = "year"))
```

Load Map Data from Shapefile: 
```{r}
setwd("C:/Users/kdfro/Documents/MSU/GitHub/ECNS-560-Term-Project")

#Load Habitats:
#habitats = raster("C:/Users/kdfro/Documents/MSU/GitHub/ECNS-560-Term-Project/Data/USGS Bison Habitat Map/mBISOx_CONUS_HabMap_2001v1/mBISOx_CONUS_HabMap_2001v1.tif")

#Load Ranges:
ranges = st_read("Data/Bison Range Map/mBISOx_CONUS_Range_2001v1/mBISOx_CONUS_Range_2001v1.shp")

```

Quick Spatial Plot:
```{r}
qtm(ranges)
qtm(habitats)
```

Additional Visualizations:

```{r}
#Comparing population change over time across some Great Plains states
graph1_data = merged_bison_data |>
  filter(state_code == "MONTANA" | state_code == "WYOMING" | state_code == "NORTH DAKOTA" | state_code == "SOUTH DAKOTA")

graph2_data = merged_bison_data |>
  filter(state_code == "MONTANA" | state_code == "WYOMING" | state_code == "NORTH DAKOTA" | state_code == "SOUTH DAKOTA" | state_code == "NEBRASKA" | state_code == "KANSAS" | state_code == "OKLAHOMA" | state_code == "TEXAS" | state_code == "IDAHO" | state_code == "COLORADO" | state_code == "NEW MEXICO")

graph3_data = merged_bison_data |>
  filter(state_code == "TEXAS")

#Some major state populations... all trends look different
ggplot(graph1_data, aes(x = year, y = total_population)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Bison Population by State - Some Major Locations", x = "Year", y = "Bison Population") +
  facet_wrap(~state_code)

#Overall trends across the great plains... see downward trend overall
ggplot(graph2_data, aes(x = year, y = total_population)) + 
  xlab("Year") +
  ylab("Population") +
  labs(title = "Bison Population Over Time") +
  geom_point(aes(color = state_code), size = 3) +
  geom_smooth()

#Number of Hunting licenses Over Time
ggplot(graph2_data, aes(x = year, y = `Paid Hunting License Holders`)) + 
  xlab("Year") +
  ylab("Population") +
  labs(title = "Number of Licenses Over Time") +
  geom_point(aes(color = state_code), size = 3) +
  geom_smooth()

#Looking for relationship between number of hunting licenses and population
ggplot(merged_bison_data, aes(x = `Paid Hunting License Holders`, y = total_population)) +
  geom_point() + 
  geom_smooth() +
  xlab("Number of Lisense Holders") +
  ylab("Population") +
  labs(title = "Number of Licenses vs Population")
```

Saki's Plots:
```{r}
data = hunting_data_tidy

names(data)[2] <- "Year"

data <- data %>%
  mutate(Year = as.numeric(Year))

data_long <- data |>
  pivot_longer(
    cols = -c(Year, state_code),  # Exclude Year and state_code from pivoting
    names_to = "License_Type",     # New column to hold the names of original columns
    values_to = "Count"            # New column to hold the values
  )

# to line plot of each license type over time for a specific state (e.g., Montana)
ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_line() +
  geom_point() +
  labs(title = "Hunting License Trends Over Time in Montana", 
       x = "Year", 
       y = "Count / Cost",
       color = "License Type") +
  theme_minimal()

ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_line() +
  geom_point() +
  labs(title = "Hunting License Trends Over Time in Montana", 
       x = "Year", 
       y = "Count / Cost",
       color = "License Type") +
  theme_minimal() +
  facet_wrap(~ License_Type, scales = "free_y")  # To separate plots by License Type with individual y-scales

ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(title = "Hunting License Trends Over Time in Montana", 
       x = "Year", 
       y = "Count / Cost",
       color = "License Type") +
  theme_minimal()

# Plot the data
ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_line() +
  geom_point() +
  scale_y_log10() +
  labs(title = "Hunting License Trends Over Time in Montana (Log Scale)", 
       x = "Year", 
       y = "Log Count / Cost",
       color = "License Type") +
  theme_minimal()

data_long$License_Type <- factor(data_long$License_Type, 
                                 labels = c("Cost Non-Resident", "Cost Resident", "Gross Cost", 
                                            "Non-Resident Licenses", "Paid Holders", "Resident Licenses"))

ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count)) +
  geom_line(aes(color = License_Type)) +
  geom_point(aes(color = License_Type)) +
  scale_y_log10() +
  labs(title = "Hunting License Trends Over Time in Montana (Log Scale)", 
       x = "Year", 
       y = "Log Count / Cost") +
  theme_minimal() +
  facet_wrap(~ License_Type, scales = "free_y")

ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_point() +
  geom_smooth(se = FALSE, method = "loess") +
  scale_y_log10() +
  labs(title = "Hunting License Trends Over Time in Montana (Log Scale)", 
       x = "Year", 
       y = "Log Count / Cost",
       color = "License Type") +
  theme_minimal()

# Load the scales package for custom label formatting
library(scales)

# Plot with custom y-axis labels
ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_point() +
  geom_smooth(se = FALSE, method = "loess") +
  scale_y_log10(labels = comma) +  # Use comma format for log scale labels
  labs(title = "Hunting License Trends Over Time in Montana (Log Scale)", 
       x = "Year", 
       y = "Count / Cost",
       color = "License Type") +
  theme_minimal()
```