### 2024-12-02 Data_Cleaning_20241031.Rmd
Add files via upload

<<<<<<< HEAD
---
title: "ECNS 560 Project - Bison Population"
author: "Kyle Froisland"
date: "2024-10-16"
output: html_document
---

Load Raw Data/Packages: *Data loaded from path on my local computer
```{r}
library(readr)
library(tidyverse)
library(readxl)
library(sf)
library(tmap)
library(raster)
library(rvest)
library(units)
library(dplyr)
library(ggplot2)
library(units)

Bison_numbers <- read_csv("C:/Users/kdfro/Documents/MSU/GitHub/ECNS-560-Term-Project/Data/Bison_numbers_by_the_county National Bison Association.csv")
hunting_data_tidy <- read_excel("C:/Users/kdfro/Documents/MSU/GitHub/ECNS-560-Term-Project/Data/hunting_data_tidy.xlsx")
```

Create Tidy Version of Population Data:
```{r}
tidy_bison = Bison_numbers |>
  pivot_longer(cols=bison_census2002:bison_census2017, names_to="year", values_to="population") |>
  dplyr::select(state_code, county, year, population)
```

Cleaning:
```{r}
#Change year Variable in tidy_bison to just the number (remove "bison_census")
tidy_bison_clean = mutate(tidy_bison, year = sapply(str_remove(tidy_bison$year, "bison_census"), function(x) (x[1])))

#define new dataset with by-state population vs by county
bison_data_merge = tidy_bison_clean |>
  group_by(state_code, year) |>
  summarize(total_population = sum(population)) |>
  rename_at(2, ~'year')

#Change column name of year in hunting data, filter to only years that match population data
colnames(hunting_data_tidy)[2] = 'year' 
hunting_data_tidy |>
  filter(year == "2002" | year == "2007" | year == "2012" | year == "2017")

#Ensure that years are numeric
bison_data_merge$year = as.numeric(bison_data_merge$year)
hunting_data_tidy$year = as.numeric(hunting_data_tidy$year)

```

Merge Data Sets:
```{r}
#join with hunting data by state and year as ID
merged_bison_data = left_join(bison_data_merge, hunting_data_tidy, by = c("state_code" = "state_code", "year" = "year"))
```

Load Map Data from Shapefile: 
```{r}
setwd("C:/Users/kdfro/Documents/MSU/GitHub/ECNS-560-Term-Project")

#Load Ranges:
ranges = st_read("Data/Bison Range Map/mBISOx_CONUS_Range_2001v1/mBISOx_CONUS_Range_2001v1.shp")

#Load States:
states = st_read("C:/Users/kdfro/Documents/MSU/GitHub/course-materials/lecture-slides/12-Spatial/data/us_states_contiguous/states_contiguous.shp")
```
Fix States and Ranges:
```{r}
#Check CRS
st_crs(states)
st_crs(ranges)

#Transform/Project:
ranges2 = st_transform(ranges, crs = st_crs(states))
stopifnot(st_crs(ranges2) == (st_crs(states)))

tm_shape(states) + tm_polygons() +
  tm_shape(ranges2) + tm_polygons(col = "darkgreen")
```

Spatial Cleaning/Exploration:
```{r}
#Split ranges df into multiple polygons
ranges3 <- st_cast(ranges2, "POLYGON")

#Create variable for area
ranges3$area = st_area(ranges3)

#Change units from m^2 to km^2
set_units(st_area(ranges3), km^2)

#Visualize
tm_shape(ranges3) + tm_polygons(col = "area")

#Select states of interest:
states2 = states |>
  filter(STATE == "Colorado" | STATE == "Idaho" | STATE == "Kansas" | STATE == "Montana" | STATE == "Nebraska" | STATE == "New Mexico" | STATE == "North Dakota" | STATE == "North Dakota" | STATE == "Oklahoma" | STATE == "South Dakota" | STATE == "Texas" | STATE == "Wyoming")

#Find intersection of ranges and states:
ranges_intersect = st_filter(ranges3, states2)

#Visualize:
tm_shape(states2) + tm_polygons() +
  tm_shape(ranges_intersect) + tm_polygons(col = "darkgreen")
```

Additional Visualizations:

```{r}
#Comparing population change over time across some Great Plains states
graph1_data = merged_bison_data |>
  filter(state_code == "MONTANA" | state_code == "WYOMING" | state_code == "NORTH DAKOTA" | state_code == "SOUTH DAKOTA")

graph2_data = merged_bison_data |>
  filter(state_code == "MONTANA" | state_code == "WYOMING" | state_code == "NORTH DAKOTA" | state_code == "SOUTH DAKOTA" | state_code == "NEBRASKA" | state_code == "KANSAS" | state_code == "OKLAHOMA" | state_code == "IDAHO" | state_code == "COLORADO" | state_code == "NEW MEXICO") |>
  mutate(population_bin = cut(total_population, breaks = c(0, 2000, 5000, 10000), labels = c("0-2000", "2000-5000", "5000-10000")))

graph3_data = merged_bison_data |>
  filter(state_code == "TEXAS")

graph4_data = merged_bison_data |>
  filter(state_code == "OKLAHOMA")

graph5_data = merged_bison_data |>
  filter(state_code == "MONTANA" | state_code == "TEXAS" | state_code == "OKLAHOMA" | state_code == "WYOMING" | state_code == "NORTH DAKOTA" | state_code == "SOUTH DAKOTA" | state_code == "NEBRASKA" | state_code == "KANSAS" | state_code == "OKLAHOMA" | state_code == "IDAHO" | state_code == "COLORADO" | state_code == "NEW MEXICO")

#Some major state populations... all trends look different
ggplot(graph1_data, aes(x = year, y = total_population)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Bison Population by State - Some Major Locations", x = "Year", y = "Bison Population") +
  facet_wrap(~state_code)

#Populations for all states of interest. There seems to be larger decreases in population around western planes states than southwestern states.
ggplot(graph5_data, aes(x = year, y = total_population)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Bison Population by State - All States of Interest", x = "Year", y = "Bison Population") +
  facet_wrap(~state_code)
#This is also biased because southwestern states started with much lower populations to begin with.

#Overall trends across the great plains... see downward trend overall
ggplot(graph2_data, aes(x = year, y = total_population)) + 
  xlab("Year") +
  ylab("Population by State") +
  labs(title = "Bison Population Over Time") +
  geom_point(aes(color = population_bin), size = 3) +
  geom_smooth()
#Each point on this plot represents a state for a given year, and the color represents the population bin that the state falls into.

#Number of Hunting licenses Over Time
ggplot(graph2_data, aes(x = year, y = `Paid Hunting License Holders`)) + 
  xlab("Year") +
  ylab("Number of Licenses") +
  labs(title = "Number of Licenses Over Time") +
  geom_point(aes(color = state_code), size = 3)
#Removed Texas from this plot because it was an extreme value and make the graph less interpretable by extending the y-axis so much. Texas saw a small increase over the last time period but was otherwise steady.

#Looking for relationship between number of hunting licenses and population
ggplot(merged_bison_data, aes(x = `Paid Hunting License Holders`, y = total_population)) +
  geom_point() + 
  geom_smooth() +
  xlab("Number of Lisense Holders") +
  ylab("Bison Population") +
  labs(title = "Number of Licenses vs Population")
#Not a clear relationship. With so many states there are too many competing factors at play to control for with our limited data.

#What about just Oklahoma since it is one of the only states with increasing numbers of licenses?
ggplot(graph4_data, aes(x = `Paid Hunting License Holders`, y = total_population)) +
  geom_point() + 
  xlab("Number of Lisense Holders") +
  ylab("Bison Population") +
  labs(title = "Number of Licenses vs Population for Oklahoma")
#We still have not inferred causality, but there seems to be a negative relationship now. The lack of variation in hunting licenses across other states makes regression extremely difficult, but Oklahoma could be a better case study.

```

Research Question 1: Can we predict percent decrease in population levels based on Latitude of their centroid (Proxied by proximity to Saskatchewan)?
```{r}
#Load Canada Shapefile:
Canada = st_read("C:/Users/kdfro/Documents/MSU/GitHub/ECNS-560-Term-Project/Data/ca_shp/ca.shp")

#Convert GRS:
Canada2 = st_transform(Canada, crs = st_crs(states2))
stopifnot(st_crs(Canada2) == (st_crs(states2)))

#Find Centroids of states of interest:
state_centroids = st_centroid(states2)

#Plot to confirm:
tm_shape(states2) + tm_borders() +
  tm_shape(states2) + tm_dots(col = "red")

#Make a variable for proximity of centroid to Saskatchewan:
distance_pop_data = state_centroids |>
  mutate(distance = st_distance(state_centroids$geometry, Canada2[3,4]))

#Merge data with population, create variables for distance from Canada and population decline:
distance = as.data.frame(distance_pop_data) |>
  rename(state_code = STATE) |>
  mutate(state_code = toupper(state_code)) |>
  left_join(bison_data_merge, by = "state_code") |>
  group_by(state_code) |>
  mutate(difference = abs(total_population - total_population[year == 2002])/total_population[year == 2002]) |>
  filter(year == 2017)

#Plot Proximity vs Population Decrease:
ggplot(distance, aes(x = log(distance), y = difference)) + 
  xlab("log(distance) from Canada") +
  ylab("Percent Decrease in Bison Population") +
  labs(title = "Latitude and Bison Population Decline") +
  geom_point()
#There does not seem to be as clear of a relationship from this plot as I had originally expected. The predictive power of this variable seems tenuous.
```

Research Question 2: Can we predict population decline using Latitude and hunting data? This probably won't work due to limited data points...
```{r}
library(tidyverse)
library(tidymodels)
library(ISLR)
library(glmnet)

#Merge Distance data set with hunting data:
distance_hunting = left_join(distance, hunting_data_tidy, by = c("state_code" = "state_code", "year" = "year")) |> as.data.frame()

#Create Predictive Model:
#First, split into train and test data
set.seed(123)
bison_split = distance_hunting |> initial_split(prop = 0.7)
bison_train = distance_hunting |> training()
bison_test = distance_hunting |> testing()

#Define a Recipe:
bison_recipe = recipe(difference ~ disatnce + total_population + Paid Hunting License Holders, data = bison_train) |> 
  update_role(GEOID, new_role = "id") |> 
  step_poly(all_numeric_predictors(), degree = 2) |> #create second term for each numeric var that is a quadratic
  step_interact(~all_numeric_predictors():all_nominal_predictors()) |> #create interactions between numeric and indicator variables
  step_dummy(all_nominal_predictors()) |> #create indicators for each categorical or character variable
  step_nzv(all_predictors())#remove predictors with nearly 0 variance... otherwise not very reliable because difficult to pickup patterns

#Process the Data:
bison_clean = bison_recipe |> prep() |> juice()

#Define linear model:
model_lm = linear_reg() |> set_engine("lm")

#Define cross validation for experimentation with other regression forms later
bison_cv = bison_train |> vfold_cv(v=5)

#Define Workflow:
fit_lm_cv = workflow() |>
  add_model(model_lm) |>
  add_recipe(bisont_recipe) |>
  fit_resamples(bison_cv)

#Fit Model:
fit_lm_cv |> collect_metrics(summarize = F)

```

Saki's Plots:
```{r}
data = hunting_data_tidy

names(data)[2] <- "Year"

data <- data %>%
  mutate(Year = as.numeric(Year))

data_long <- data |>
  pivot_longer(
    cols = -c(Year, state_code),  # Exclude Year and state_code from pivoting
    names_to = "License_Type",     # New column to hold the names of original columns
    values_to = "Count"            # New column to hold the values
  )

# to line plot of each license type over time for a specific state (e.g., Montana)
ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_line() +
  geom_point() +
  labs(title = "Hunting License Trends Over Time in Montana", 
       x = "Year", 
       y = "Count / Cost",
       color = "License Type") +
  theme_minimal()

ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_line() +
  geom_point() +
  labs(title = "Hunting License Trends Over Time in Montana", 
       x = "Year", 
       y = "Count / Cost",
       color = "License Type") +
  theme_minimal() +
  facet_wrap(~ License_Type, scales = "free_y")  # To separate plots by License Type with individual y-scales

ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(title = "Hunting License Trends Over Time in Montana", 
       x = "Year", 
       y = "Count / Cost",
       color = "License Type") +
  theme_minimal()

# Plot the data
ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_line() +
  geom_point() +
  scale_y_log10() +
  labs(title = "Hunting License Trends Over Time in Montana (Log Scale)", 
       x = "Year", 
       y = "Log Count / Cost",
       color = "License Type") +
  theme_minimal()

data_long$License_Type <- factor(data_long$License_Type, 
                                 labels = c("Cost Non-Resident", "Cost Resident", "Gross Cost", 
                                            "Non-Resident Licenses", "Paid Holders", "Resident Licenses"))

ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count)) +
  geom_line(aes(color = License_Type)) +
  geom_point(aes(color = License_Type)) +
  scale_y_log10() +
  labs(title = "Hunting License Trends Over Time in Montana (Log Scale)", 
       x = "Year", 
       y = "Log Count / Cost") +
  theme_minimal() +
  facet_wrap(~ License_Type, scales = "free_y")

ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_point() +
  geom_smooth(se = FALSE, method = "loess") +
  scale_y_log10() +
  labs(title = "Hunting License Trends Over Time in Montana (Log Scale)", 
       x = "Year", 
       y = "Log Count / Cost",
       color = "License Type") +
  theme_minimal()

# Load the scales package for custom label formatting
library(scales)

# Plot with custom y-axis labels
ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_point() +
  geom_smooth(se = FALSE, method = "loess") +
  scale_y_log10(labels = comma) +  # Use comma format for log scale labels
  labs(title = "Hunting License Trends Over Time in Montana (Log Scale)", 
       x = "Year", 
       y = "Count / Cost",
       color = "License Type") +
  theme_minimal()
```
=======
---
title: "ECNS 560 Project - Bison Population"
author: "Kyle Froisland"
date: "2024-10-16"
output: html_document
---

Load Raw Data/Packages: *Data loaded from path on my local computer
```{r}
library(readr)
library(tidyverse)
library(readxl)
library(sf)
library(tmap)
library(raster)
library(rvest)
library(units)
library(dplyr)
library(ggplot2)
library(units)

Bison_numbers <- read_csv("C:/Users/kdfro/Documents/MSU/GitHub/ECNS-560-Term-Project/Data/Bison_numbers_by_the_county National Bison Association.csv")
hunting_data_tidy <- read_excel("C:/Users/kdfro/Documents/MSU/GitHub/ECNS-560-Term-Project/Data/hunting_data_tidy.xlsx")
```

Create Tidy Version of Population Data:
```{r}
tidy_bison = Bison_numbers |>
  pivot_longer(cols=bison_census2002:bison_census2017, names_to="year", values_to="population") |>
  dplyr::select(state_code, county, year, population)
```

Cleaning:
```{r}
#Change year Variable in tidy_bison to just the number (remove "bison_census")
tidy_bison_clean = mutate(tidy_bison, year = sapply(str_remove(tidy_bison$year, "bison_census"), function(x) (x[1])))

#define new dataset with by-state population vs by county
bison_data_merge = tidy_bison_clean |>
  group_by(state_code, year) |>
  summarize(total_population = sum(population)) |>
  rename_at(2, ~'year')

#Change column name of year in hunting data, filter to only years that match population data
colnames(hunting_data_tidy)[2] = 'year' 
hunting_data_tidy |>
  filter(year == "2002" | year == "2007" | year == "2012" | year == "2017")

#Ensure that years are numeric
bison_data_merge$year = as.numeric(bison_data_merge$year)
hunting_data_tidy$year = as.numeric(hunting_data_tidy$year)

```

Merge Data Sets:
```{r}
#join with hunting data by state and year as ID
merged_bison_data = left_join(bison_data_merge, hunting_data_tidy, by = c("state_code" = "state_code", "year" = "year"))
```

Load Map Data from Shapefile: 
```{r}
setwd("C:/Users/kdfro/Documents/MSU/GitHub/ECNS-560-Term-Project")

#Load Ranges:
ranges = st_read("Data/Bison Range Map/mBISOx_CONUS_Range_2001v1/mBISOx_CONUS_Range_2001v1.shp")

#Load States:
states = st_read("C:/Users/kdfro/Documents/MSU/GitHub/course-materials/lecture-slides/12-Spatial/data/us_states_contiguous/states_contiguous.shp")
```
Merge States and Ranges:
```{r}
#Check CRS
st_crs(states)
st_crs(ranges)

#Transform/Project:
ranges2 = st_transform(ranges, crs = st_crs(states))
stopifnot(st_crs(ranges2) == (st_crs(states)))

tm_shape(states) + tm_polygons() +
  tm_shape(ranges2) + tm_polygons(col = "darkgreen")
```

Spatial Cleaning/Exploration:
```{r}
#Split ranges df into multiple polygons
ranges3 <- st_cast(ranges2, "POLYGON")

#Create variable for area
ranges3$area = st_area(ranges3)

#Change units from m^2 to km^2
set_units(st_area(ranges3), km^2)

#Visualize
tm_shape(ranges3) + tm_polygons(col = "area")

#Select states of interest:
states2 = states |>
  filter(STATE == "Colorado" | STATE == "Idaho" | STATE == "Kansas" | STATE == "Montana" | STATE == "Nebraska" | STATE == "New Mexico" | STATE == "North Dakota" | STATE == "North Dakota" | STATE == "Oklahoma" | STATE == "South Dakota" | STATE == "Texas" | STATE == "Wyoming")

#Find intersection of ranges and states:
ranges_intersect = st_filter(ranges3, states2)

#Visualize:
tm_shape(states2) + tm_polygons() +
  tm_shape(ranges_intersect) + tm_polygons(col = "darkgreen")
```

Additional Visualizations:

```{r}
#Comparing population change over time across some Great Plains states
graph1_data = merged_bison_data |>
  filter(state_code == "MONTANA" | state_code == "WYOMING" | state_code == "NORTH DAKOTA" | state_code == "SOUTH DAKOTA")

graph2_data = merged_bison_data |>
  filter(state_code == "MONTANA" | state_code == "WYOMING" | state_code == "NORTH DAKOTA" | state_code == "SOUTH DAKOTA" | state_code == "NEBRASKA" | state_code == "KANSAS" | state_code == "OKLAHOMA" | state_code == "IDAHO" | state_code == "COLORADO" | state_code == "NEW MEXICO") |>
  mutate(population_bin = cut(total_population, breaks = c(0, 2000, 5000, 10000), labels = c("0-2000", "2000-5000", "5000-10000")))

graph3_data = merged_bison_data |>
  filter(state_code == "TEXAS")

graph4_data = merged_bison_data |>
  filter(state_code == "OKLAHOMA")

#Some major state populations... all trends look different
ggplot(graph1_data, aes(x = year, y = total_population)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Bison Population by State - Some Major Locations", x = "Year", y = "Bison Population") +
  facet_wrap(~state_code)

#Overall trends across the great plains... see downward trend overall
ggplot(graph2_data, aes(x = year, y = total_population)) + 
  xlab("Year") +
  ylab("Population by State") +
  labs(title = "Bison Population Over Time") +
  geom_point(aes(color = population_bin), size = 3) +
  geom_smooth()
#Each point on this plot represents a state for a given year, and the color represents the population bin that the state falls into.

#Number of Hunting licenses Over Time
ggplot(graph2_data, aes(x = year, y = `Paid Hunting License Holders`)) + 
  xlab("Year") +
  ylab("Number of Licenses") +
  labs(title = "Number of Licenses Over Time") +
  geom_point(aes(color = state_code), size = 3)
#Removed Texas from this plot because it was an extreme value and make the graph less interpretable by extending the y-axis so much. Texas saw a small increase over the last time period but was otherwise steady.

#Looking for relationship between number of hunting licenses and population
ggplot(merged_bison_data, aes(x = `Paid Hunting License Holders`, y = total_population)) +
  geom_point() + 
  geom_smooth() +
  xlab("Number of Lisense Holders") +
  ylab("Bison Population") +
  labs(title = "Number of Licenses vs Population")
#Not a clear relationship. With so many states there are too many competing factors at play to control for with our limited data.

#What about just Oklahoma since it is one of the only states with increasing numbers of licenses?
ggplot(graph4_data, aes(x = `Paid Hunting License Holders`, y = total_population)) +
  geom_point() + 
  xlab("Number of Lisense Holders") +
  ylab("Bison Population") +
  labs(title = "Number of Licenses vs Population for Oklahoma")
#We still have not inferred causality, but there seems to be a negative relationship now. The lack of variation in hunting licenses across other states makes regression extremely difficult, but Oklahoma could be a better case study.

```

Saki's Plots:
```{r}
data = hunting_data_tidy

names(data)[2] <- "Year"

data <- data %>%
  mutate(Year = as.numeric(Year))

data_long <- data |>
  pivot_longer(
    cols = -c(Year, state_code),  # Exclude Year and state_code from pivoting
    names_to = "License_Type",     # New column to hold the names of original columns
    values_to = "Count"            # New column to hold the values
  )

# to line plot of each license type over time for a specific state (e.g., Montana)
ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_line() +
  geom_point() +
  labs(title = "Hunting License Trends Over Time in Montana", 
       x = "Year", 
       y = "Count / Cost",
       color = "License Type") +
  theme_minimal()

ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_line() +
  geom_point() +
  labs(title = "Hunting License Trends Over Time in Montana", 
       x = "Year", 
       y = "Count / Cost",
       color = "License Type") +
  theme_minimal() +
  facet_wrap(~ License_Type, scales = "free_y")  # To separate plots by License Type with individual y-scales

ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(title = "Hunting License Trends Over Time in Montana", 
       x = "Year", 
       y = "Count / Cost",
       color = "License Type") +
  theme_minimal()

# Plot the data
ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_line() +
  geom_point() +
  scale_y_log10() +
  labs(title = "Hunting License Trends Over Time in Montana (Log Scale)", 
       x = "Year", 
       y = "Log Count / Cost",
       color = "License Type") +
  theme_minimal()

data_long$License_Type <- factor(data_long$License_Type, 
                                 labels = c("Cost Non-Resident", "Cost Resident", "Gross Cost", 
                                            "Non-Resident Licenses", "Paid Holders", "Resident Licenses"))

ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count)) +
  geom_line(aes(color = License_Type)) +
  geom_point(aes(color = License_Type)) +
  scale_y_log10() +
  labs(title = "Hunting License Trends Over Time in Montana (Log Scale)", 
       x = "Year", 
       y = "Log Count / Cost") +
  theme_minimal() +
  facet_wrap(~ License_Type, scales = "free_y")

ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_point() +
  geom_smooth(se = FALSE, method = "loess") +
  scale_y_log10() +
  labs(title = "Hunting License Trends Over Time in Montana (Log Scale)", 
       x = "Year", 
       y = "Log Count / Cost",
       color = "License Type") +
  theme_minimal()

# Load the scales package for custom label formatting
library(scales)

# Plot with custom y-axis labels
ggplot(data_long %>% filter(state_code == "MONTANA"), aes(x = Year, y = Count, color = License_Type)) +
  geom_point() +
  geom_smooth(se = FALSE, method = "loess") +
  scale_y_log10(labels = comma) +  # Use comma format for log scale labels
  labs(title = "Hunting License Trends Over Time in Montana (Log Scale)", 
       x = "Year", 
       y = "Count / Cost",
       color = "License Type") +
  theme_minimal()
```
>>>>>>> ae5d3f30ef6b3c153d8c3a489f3f9dc3d1e461cb